---
title: "Hands-on Exercise 10"
author: "Brigitta Karen Tsai"
date: "October 31, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

# Spatial Interaction Models

# 10A: Processing and Visualising Flow Data

## 1 Getting Started

```{r}
pacman::p_load(tmap, sf, DT, stplanr, tidyverse)
```

## 2 Preparing the Flow Data

### 2.1 Importing OD Data

Import Passenger Volume by Origin Destination Bus Stops

```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202210.csv")
```

```{r}
glimpse(odbus)
```

```{r}
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

### 2.2 Extracting the Study Data

```{r}
odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
datatable(odbus6_9)
```

```{r}
write_rds(odbus6_9, "data/rds/odbus6_9.rds")
```

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```

## 3 Working with Geospatial Data

### 3.1 Import Geospatial Data

```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz
```

```{r}
mpsz <- write_rds(mpsz, "data/rds/mpsz.rds")
```

## 4 Geospatial Data Wrangling

### 4.1 Combining Bus Stop and MPSZ

```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```

```{r}
datatable(busstop_mpsz)
```

```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.rds")  
```

Append planning subzone code from busstop_mpsz data onto odbus6_9 data

```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```

```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
od_data <- unique(od_data)
```

```{r}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(MORNING_PEAK = sum(TRIPS))
```

```{r}
write_rds(od_data, "data/rds/od_data_fii.rds")
```

```{r}
od_data_fii <- read_rds("data/rds/od_data.rds")
```

## 5 Visualising Spatial Interaction

### 5.1 Removing Intra-zonal Flows

```{r}
od_data_fij <- od_data[od_data$ORIGIN_SZ!=od_data$DESTIN_SZ,]
```

```{r}
write_rds(od_data_fij, "data/rds/od_data_fij.rds")
```

```{r}
od_data_fij <- read_rds("data/rds/od_data_fij.rds")
```

### 5.2 Creating Desire Lines

```{r}
flowLine <- od2line(flow = od_data_fij, 
                    zones = mpsz,
                    zone_code = "SUBZONE_C")
```

```{r}
write_rds(flowLine, "data/rds/flowLine.rds")
```

```{r}
flowLine <- read_rds("data/rds/flowLine.rds")
```

### 5.3 Visualising the Desire Lines

```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```

```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
  filter(MORNING_PEAK >= 5000) %>%
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```

# 10B: Calibrating Spatial Interaction Models

## 1 Getting Started

```{r}
pacman::p_load(tmap, sf, sp,
               performance, reshape2,
               ggpubr, tidyverse)
```

## 2 Computing Distance Matrix

```{r}
mpsz
```

### 2.1 Convert sf to Spatial Polygon Dataframe

```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

### 2.2 Computing the distance matrix

```{r}
dist <- spDists(mpsz_sp, 
                longlat = FALSE)
```

```{r}
head(dist, n=c(10, 10))
```

### 2.3 Labelling Column and Row Headers of a distance matrix

```{r}
sz_names <- mpsz$SUBZONE_C
```

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```

### 2.4 Pivoting Distance Value by Subzone_C

```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```

### 2.5 Updating Intra-zonal Distances

```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()
```

```{r}
distPair$dist <- ifelse(distPair$dist == 0,
                        50, distPair$dist)
```

```{r}
distPair %>%
  summary()
```

```{r}
distPair <- distPair %>%
  rename(orig = Var1,
         dest = Var2)
```

```{r}
write_rds(distPair, "data/rds/distPair.rds") 
```

```{r}
distPair <- read_rds("data/rds/distPair.rds")
```

## 3 Preparing the Flow Data

```{r}
od_data_fii <- read_rds("data/rds/od_data_fii.rds")
```

```{r}
flow_data <- od_data_fii %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>% 
  summarize(TRIPS = sum(MORNING_PEAK)) 
```

```{r}
head(flow_data, 10)
```

### 3.1 Separating Intra-flow from passenger volume df

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0.000001, 1)
```

### 3.2 Combining Passenger Volume Data with Distance Value

```{r}
flow_data$ORIGIN_SZ <- as.factor(flow_data$ORIGIN_SZ)
flow_data$DESTIN_SZ <- as.factor(flow_data$DESTIN_SZ)
```

```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_SZ" = "orig",
                    "DESTIN_SZ" = "dest"))
```

## 4 Preparing Origin and Destination Attributes

### 4.1 Importing Population Data

```{r}
pop <- read_csv("data/aspatial/pop.csv")
```

### 4.2 Geospatial Data Wrangling

```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA" = "PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)
```

### 4.3 Preparing Origin Attribute

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

### 4.4 Preparing Destination Attribute

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

```{r}
write_rds(flow_data1, "data/rds/SIM_data.rds")
```

## 5 Calibrating Spatial Interaction Models

### 5.1 Importing the Modelling Data

```{r}
SIM_data <- read_rds("data/rds/SIM_data.rds")
```

### 5.2 Visualising the Dependent Variable

```{r}
ggplot(data = SIM_data,
       aes(x = TRIPS)) +
  geom_histogram()
```

```{r}
ggplot(data = SIM_data,
       aes(x = dist,
           y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ggplot(data = SIM_data,
       aes(x = log(dist),
           y = log(TRIPS))) +
  geom_point() +
  geom_smooth(method = lm)
```

### 5.3 Checking for Variables with Zero Values

```{r}
summary(SIM_data)
```

```{r}
SIM_data$DESTIN_AGE7_12 <- ifelse(
  SIM_data$DESTIN_AGE7_12 == 0,
  0.99, SIM_data$DESTIN_AGE7_12)
SIM_data$DESTIN_AGE13_24 <- ifelse(
  SIM_data$DESTIN_AGE13_24 == 0,
  0.99, SIM_data$DESTIN_AGE13_24)
SIM_data$DESTIN_AGE25_64 <- ifelse(
  SIM_data$DESTIN_AGE25_64 == 0,
  0.99, SIM_data$DESTIN_AGE25_64)
SIM_data$ORIGIN_AGE7_12 <- ifelse(
  SIM_data$ORIGIN_AGE7_12 == 0,
  0.99, SIM_data$ORIGIN_AGE7_12)
SIM_data$ORIGIN_AGE13_24 <- ifelse(
  SIM_data$ORIGIN_AGE13_24 == 0,
  0.99, SIM_data$ORIGIN_AGE13_24)
SIM_data$ORIGIN_AGE25_64 <- ifelse(
  SIM_data$ORIGIN_AGE25_64 == 0,
  0.99, SIM_data$ORIGIN_AGE25_64)
```

```{r}
summary(SIM_data)
```

### 5.4 Unconstrained Spatial Interaction Model

```{r}
uncSIM <- glm(formula = TRIPS ~ 
                log(ORIGIN_AGE25_64) + 
                log(DESTIN_AGE25_64) +
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
uncSIM
```

### 5.5 R Squared Function

```{r}
CalcRSquared <- function(observed,estimated){
  r <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

```{r}
CalcRSquared(uncSIM$data$TRIPS, uncSIM$fitted.values)
```

```{r}
r2_mcfadden(uncSIM)
```

### 5.6 Origin (Production) Constrained SIM

```{r}
orcSIM <- glm(formula = TRIPS ~ 
                 ORIGIN_SZ +
                 log(DESTIN_AGE25_64) +
                 log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(orcSIM)
```

```{r}
CalcRSquared(orcSIM$data$TRIPS, orcSIM$fitted.values)
```

### 5.7 Destination Constrained

```{r}
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_SZ + 
                log(ORIGIN_AGE25_64) + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(decSIM)
```

```{r}
CalcRSquared(decSIM$data$TRIPS, decSIM$fitted.values)
```

### 5.8 Doubly Constrained

```{r}
dbcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_SZ + 
                DESTIN_SZ + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(dbcSIM)
```

```{r}
CalcRSquared(dbcSIM$data$TRIPS, dbcSIM$fitted.values)
```

### 5.9 Model Comparison

```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)
```

```{r}
compare_performance(model_list,
                    metrics = "RMSE")
```

### 5.10 Visualising Fitted Values

```{r}
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(uncTRIPS = "uncSIM$fitted.values")
```

```{r}
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(orcTRIPS = "orcSIM$fitted.values")
```

```{r}
df <- as.data.frame(decSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(decTRIPS = "decSIM$fitted.values")
```

```{r}
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(dbcTRIPS = "dbcSIM$fitted.values")
```

```{r}
unc_p <- ggplot(data = SIM_data,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = SIM_data,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = SIM_data,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = SIM_data,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ggarrange(unc_p, orc_p, dec_p, dbc_p,
          ncol = 2,
          nrow = 2)
```
