---
title: "In-class Exercise 7"
author: "Brigitta Karen Tsai"
date: "October 14, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

## 1 Load and Install R Packages

```{r}
pacman::p_load(olsrr, corrplot, ggpubr,
               sf, spdep, GWmodel, tmap,
               tidyverse, gtsummary, ggstatsplot,
               sfdep)
```

## 2 Importing Data

```{r}
condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

## 3 Correlation Analysis

```{r}
#| fig-width: 12
#| fig-height: 10
ggcorrmat(condo_resale[, 5:23])
```

## 4 Linear Regression Method

```{r}
condo_mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM +
                  AGE + PROX_CBD + PROX_CHILDCARE +
                  PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA +
                  PROX_HAWKER_MARKET + PROX_KINDERGARTEN +
                  PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH +
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL +
                  PROX_SUPERMARKET + PROX_BUS_STOP +
                  NO_Of_UNITS + FAMILY_FRIENDLY +
                  FREEHOLD + LEASEHOLD_99YR,
                data=condo_resale.sf)
summary(condo_mlr)
```

## 5 olsrr Method

```{r}
ols_regress(condo_mlr)
```

### 5.1 Multicolinearity

```{r}
ols_vif_tol(condo_mlr)
```

### 5.2 Variable Selection

```{r}
condo_fw_mlr <- ols_step_forward_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE # TRUE = multiple iterations
)
```

```{r}
#| fig-width: 12
#| fig-height: 10
plot(condo_fw_mlr)
```

### 5.3 Test for Non-Linearity

```{r}
ols_plot_resid_fit(condo_fw_mlr$model) # no need to choose variable manually 
```

```{r}
ols_plot_resid_hist(condo_fw_mlr$model)
```

```{r}
ols_test_normality(condo_fw_mlr$model)
```

### 5.4 Testing for Spatial Autocorrelation

```{r}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals) %>% 
  rename('FW_MLR_RES' = 'condo_fw_mlr$model$residuals')
```

```{r}
condo_resale_sf <- cbind(condo_resale.sf,
                         mlr_output$FW_MLR_RES) %>% 
  rename('MLR_RES' = 'mlr_output.FW_MLR_RES')
```

```{r}
tmap_mode("plot")
tm_shape(mpsz) + 
  tmap_options(check.and.fix = TRUE) + # to fix geometric problems
  tm_polygons(alpha = 0.4) +
  tm_shape(condo_resale_sf) + 
    tm_dots(col = "MLR_RES",
            alpha = 0.6,
            style = "quantile")
```

```{r}
condo_resale_sf <- condo_resale_sf %>% 
  mutate(nb = st_knn(geometry, k = 6,
                     longlat = FALSE),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

```{r}
global_moran_perm(condo_resale_sf$MLR_RES,
                  condo_resale_sf$nb,
                  condo_resale_sf$wt,
                  alternative = "two.sided",
                  nsim = 99)
```

The global Moran's I test for residual spatial autocorrelation shows that the p-value is less than 0.00000000000000022 , which is less then the alpha of 0.05. Hence, we will reject the null hypothesis that the residuals are randomly distributed (if it's rejected, it's not random)

### 5.5 Fixed Bandwidth

```{r}
bw_fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE +
                     PROX_CBD + PROX_CHILDCARE +
                     PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA +
                     PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH +
                     PROX_SHOPPING_MALL + PROX_BUS_STOP +
                      NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                   data = condo_resale_sf,
                   approach = "CV",
                   kernel = "gaussian", # must be consistent 
                   adaptive = FALSE,
                   longlat = FALSE)
```

Now we use the code below to calibrate the gwr model using fixed bandwidth and gaussian kernel

```{r}
gwr_fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE +
                     PROX_CBD + PROX_CHILDCARE +
                     PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA +
                     PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH +
                     PROX_SHOPPING_MALL + PROX_BUS_STOP +
                      NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                   data = condo_resale_sf,
                   bw = bw_fixed,
                   kernel = "gaussian", # must be consistent 
                   adaptive = FALSE,
                   longlat = FALSE)
```

```{r}
gwr_fixed
```

### 5.6 Converting SDF into sf dataframe

```{r}
gwr_fixed_output <- as.data.frame(
  gwr_fixed$SDF) %>% 
  select(-c(2:15))
```

```{r}
gwr_sf_fixed <- cbind(condo_resale_sf,
                      gwr_fixed_output)
```

```{r}
glimpse(gwr_sf_fixed)
```
